### 2016-01-29 PPM.tex
Modified definition of deformational Courant number
Added Hilary's alternative description of PPM. Is it correct?

The dimensionally split scheme described in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian \citep{Lin1996}. This dimension splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensions.
\subsubsection{Long time-step permitting PPM} 
\label{sec:PPM}
Both PPM and the long time steps technique use subgrid cell reconstruction. Although they are defined for varying grid spacing, we will only describe them for uniform grid interval, $\Delta x$. The scheme is described on a domain with length $L$, which is divided into $n$ grid points. The position of $j$th grid is $x_j = j\Delta x$ and $j = 0, 1, 2, \cdots, n-1$ and the corresponding tracer is $\phi_j$. Subgrid cell reconstruction uses the space average mass within the $j$th grid box at time $t_n$, which is defined as: 
\begin{equation} \label{eq:2.1} 
\phi^n_j = \frac{1}{\Delta x} \int^{x_{j+\frac{1}{2}}}_{x_{j-\frac{1}{2}}} \phi (x,t)dx
\end{equation}
where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid point. Thus, $x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$.
\begin{figure}
% \vspace{-20pt}
\centering
\includegraphics[scale=0.4]{YumengsGraphics/nirvana.pdf}
% \vspace{-20pt}
\caption{The 1D long time permitting technique. The mass between $x_{j,L}$ and $x_{j-c_N-1,R}$ can be calculated using equation \ref{eq:2.3}, and the mass between $x_{j-c_N-1,R}$ and $x_{j-c}$ should be calculated using PPM \label{fig:2.1}}
\end{figure}

The long time step permitting technique is illustrated in figure \ref{fig:2.1}. Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point and the time step is restricted by the deformational Courant number $c_d$. The deformational Courant number is $\Delta t \max(|u_x|,|u_y|,|v_x|,|v_x|)$ \citep{PS84} and the sufficient condition for the stability is $c_d \leq 1$. \cite{Leonard1995} shows that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$.

It is convinient to divide the Courant number, $c$, into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$ \citep{Leonard1995}. In this way, the CFL restriction is removed, because the remnant Courant number is always less than one. The flux equation becomes:
\begin{eqnarray} \label{eq:2.2}
c\cdot \Delta x \phi_{j-\frac{1}{2}} = M_{c_r,R,j-1} + M_{j-\frac{1}{2}}-M_{j-c_N-\frac{1}{2}} & \text{c}>0 &\nonumber \\
c\cdot \Delta x \phi_{j-\frac{1}{2}} = -M_{c_r,R,j-1} +M_{j+c_N-\frac{1}{2}}- M_{j-\frac{1}{2}} & \text{c}<0 & \\
c\cdot \Delta x \phi_{j-\frac{1}{2}} = 0 & \text{otherwise} \nonumber 
\end{eqnarray}
where $M_{j-\frac{1}{2}}$ is the cumulative mass from the start point to the $j$th left zone edges defined as \citep{Colella1984}:
\begin{equation} \label{eq:2.3}
M_{j-\frac{1}{2}} = \sum_{k < j} \phi_{k}\Delta x
\end{equation}

>>>>

The one-dimensional scheme that we use is the piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995}. This is extended to two-dimensions using COSMIC splitting \citep{Leonard1996} (Conservative Operator Splitting for Multidimensions with Inherently Constancy). PPM with a long time-step is a one-dimensional flux-form semi-Lagrangian scheme \citep{Lin1996}.

\subsubsection{Long time-step permitting PPM} 
\label{sec:PPM}

Both PPM and the long time steps technique use subgrid cell reconstruction. Although they are defined for varying grid spacing, we will only describe them for uniform grid interval, $\Delta x$. We define the position of $j$th grid cell centre to be $x_j = j\Delta x$ and $j = 0, 1, 2, \cdots$ and the corresponding tracer value to be $\phi_j$. Subgrid cell reconstruction uses the space average mass within the $j$th grid box which is defined as: 
\begin{equation} \label{eq:2.1} 
\phi_j = \frac{1}{\Delta x} \int^{x_{j+\frac{1}{2}}}_{x_{j-\frac{1}{2}}} \phi (x,t)dx
\end{equation}
where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid box ($x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$).
\begin{figure}
% \vspace{-20pt}
\centering
\includegraphics[scale=0.4]{YumengsGraphics/nirvana.pdf}
% \vspace{-20pt}
\caption{The 1D long time permitting technique. The mass between $x_{j,L}$ and $x_{j-c_N-1,R}$ can be calculated using equation \ref{eq:2.3}, and the mass between $x_{j-c_N-1,R}$ and $x_{j-c}$ should be calculated using PPM \label{fig:2.1}}
\end{figure}

The long time step permitting technique is illustrated in figure \ref{fig:2.1}. Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point in one dimension. The time step is restricted by the deformational Courant number $c_d$:
\begin{equation}
c_d=
\Delta t \max
\left(\biggl|\frac{\partial u}{\partial x}\biggr|,\biggl|\frac{\partial u}{\partial y}\biggr|,\biggl|\frac{\partial u}{\partial z}\biggr|,\biggl|\frac{\partial v}{\partial x}\biggr|,\biggl|\frac{\partial v}{\partial y}\biggr|,\biggl|\frac{\partial v}{\partial z}\biggr|,\biggl|\frac{\partial w}{\partial x}\biggr|,\biggl|\frac{\partial w}{\partial y}\biggr|,\biggl|\frac{\partial w}{\partial z}\biggr|\right)
\label{eq:cd}
\end{equation}
\citep{PS84} and the stability criterion is $c_d \leq 1$.

\cite{Leonard1995} showed that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$.

It is convenient to divide the Courant number, $c$, into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$ \citep{Leonard1995}. In this way, the CFL restriction is removed, because the remnant Courant number is always less than one. The flux equation becomes:
\begin{eqnarray} \label{eq:2.2}
c \Delta x \phi_{j-\frac{1}{2}} = M_{c_r,R,j-1} + M_{j-\frac{1}{2}}-M_{j-c_N-\frac{1}{2}} & \text{c}>0 &\nonumber \\
c \Delta x \phi_{j-\frac{1}{2}} = -M_{c_r,R,j-1} +M_{j+c_N-\frac{1}{2}}- M_{j-\frac{1}{2}} & \text{c}<0 & \\
c \Delta x \phi_{j-\frac{1}{2}} = 0 & \text{otherwise} \nonumber 
\end{eqnarray}
where $M_{j-\frac{1}{2}}$ is the cumulative mass from the start point to the $j$th left zone edges defined as \citep{Colella1984}:
\begin{equation} \label{eq:2.3}
M_{j-\frac{1}{2}} = \sum_{k < j} \phi_{k}\Delta x
\end{equation}
